---
title: "Assignment 2"
author: "Chloe Moore"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
```

### Question 11. Exercise 2.20

```{r}
nursing_homeDF = read.csv("WiscNursingHome.csv")
nursing_home2001 = subset(nursing_homeDF, CRYEAR == 2001) %>%
  mutate(
    LOGTPY = log(TPY),
    LOGNUMBED = log(NUMBED)
  )
```

#### Part a)
```{r}
summary(nursing_home2001$LOGTPY) # summary statistics LOGTPY
summary(nursing_home2001$LOGNUMBED) # summary statistics LOGNUMBED

cor(nursing_home2001$LOGTPY, nursing_home2001$LOGNUMBED) # correlation statistic

ggplot(nursing_home2001, aes(x=LOGNUMBED, y=LOGTPY)) +
  geom_point()
```

\n Comments: The correlation statistic is very close to 1, which indicates LOGTPY and LOGNUMBED are closely, positively correlated. The scatter plot chart appears to show a linear trend.

#### Part b)
```{r}
fit1 = lm(LOGTPY ~ LOGNUMBED, nursing_home2001)
summary1 = summary(fit1)
summary1
```
\n Comments: R^2 = 0.9664 (adjusted R^2=0.9663). Regression coefficient LOGNUMBED = 1.01923. t-statistic = 100.73

#### Part c)
```{r}
b1 = summary1$coefficients[2, 1]
seb1 = summary1$coefficients[2, 2]
tcrit = 2 # for t_(353, 0.975)
```

\n (I) $H_0:\beta_{1}=0$ vs $H_a:\beta_{1}\ne0$. We reject the null hypothesis because our t-statistic is greater than the critical t-value. The p-value is less than our alpha level of 0.05, which supports this rejection.

```{r}
tstat = b1/seb1 # 100.730193
pval = 2*(1-pt(abs(tstat), 353)) # 0.0000000...
```

\n (II) $H_0:\beta_{1}=1$ vs $H_a:\beta_{1}\ne1$. We fail to reject the null hypothesis because our t-statistic is less than the critical t-value (however, it's very close!). The p-value is greater than our alpha level of 0.05, which supports this conclusion.

```{r}
tstat = (b1-1)/seb1 # 1.900567
pval = 2*(1-pt(abs(tstat), 353)) # 0.058173
```

\n (III) $H_0:\beta_{1}=1$ vs $H_a:\beta_{1}>1$. We might fail to reject the null hypothesis because our t-statistic is less than the critical t-value... But the p-value is less than our alpha level of 0.05, so in this case, we will actually reject the null hypothesis.

```{r}
tstat = (b1-1)/seb1 # 1.900567
pval = (1-pt(abs(tstat), 353)) # 0.029087
```

\n (IV) $H_0:\beta_{1}=1$ vs $H_a:\beta_{1}<1$. We might fail to reject the null hypothesis because our t-statistic is less than the critical t-value... And the p-value is greater than our alpha level of 0.05, so we will indeed fail to reject the null hypothesis.

```{r}
tstat = (b1-1)/seb1 # 1.900567
pval = pt(abs(tstat), 353) # 0.970913
```






